Modeling the impacts of park access on health outcomes: a utility-based accessibility approach

### 2020-06-04 manuscript/submission/manuscript.tex
Fix affiliations and update title page

\documentclass[Afour,sageh.bst]{sagej}

@@

\usepackage{bm}
     \def\Var{{\rm Var}\,}
     \def\Cov{{\rm Cov}\,}
     \def\mean{{\rm mean}\,}
     \def\E{{\rm E}\,}
  	 \def\Corr{{\rm Corr}\,}
\newcommand{\vect}[1]{\boldsymbol{#1}}

\usepackage{hyperref}
\usepackage{rotating}
\usepackage{booktabs}
\usepackage{makecell}

@@

\runninghead{XXXX \emph{et al}.}

\author{XXXX \affilnum{1}, XXXX \affilnum{2}, XXXX\affilnum{3}, XXXX\affilnum{3}}

\affiliation{
\affilnum{1}{XXXX}\\
\affilnum{2}{XXXX}\\
\affilnum{3}{XXXX}}

\corrauth{XXXX}

\email{\href{XXXX}{\nolinkurl{XXXX}}}

@@

\keywords{Accessibility; Green Space; Public Health}

\maketitle

\clearpage
\newpage
\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

@@

Historically, most large-scale health promotion efforts focused on
individual-level interventions intended to educate people about healthy
lifestyles and behaviors, touching on topics including diet and
exercise. Over the last several years a new \emph{social} model of
health has evolved, described by \citet{Duhl1999} ``as an outcome of the
effects of socioeconomic status, culture, environmental conditions,
housing, employment and community
influences''\citeyearpar[p.7]{Duhl1999}. In this new paradigm, resources
provided via civil infrastructure --- in particular parks and other
public green spaces --- play a crucial role in promoting and sustaining
public health \citep{Bedimo-Rung2005, Wells2007, Coutts2008}. After all,
it does little to educate individuals on the importance of exercise if
their community is built in such a way that such exercise is expensive,
unenjoyable, or unsafe.

In spite of the potential for green spaces to influence public health
outcomes and numerous studies attempting to empirically estimate the
strength of these outcomes, evidence to date has been mixed. A major
challenge researchers have faced is a proliferation of techniques for
measuring access to green spaces, of which many are plausible but lack
robust theoretical bases. In this paper, we apply a holistic and
flexible measurement for park accessibility based in activity location
choice theory. This utility-based accessibility measure compares the
continuous distance to all parks in the region, weighted against the
sizes of each park and its other assorted amenities.

We apply this measurement to study the link between park accessibility
and attractiveness and census tract-level aggregate physical activity
participation and obesity rates in New York City, controlling for
spatially correlated unobserved effects, spatial spillovers of
covariates, and tract-level socioeconomic characteristics. We also
demonstrate how the utility-based accessibility metric can potentially
be expanded to account for other measures of a park's attractiveness
including its observed social media activity and the presence of various
park amenities. The paper concludes with a discussion of opportunities
for future research.

@@

\label{sec:litreview}

The large majority of extant research on the topic of park access and
public health outcomes has been focused on comparing metropolitan
regions against each other. For example, \citet{West2012} used park data
from the Trust for Public Land's 2010 ``City Park Facts'' and public
health data from the Behavioral Risk Factor Surveillance System (BRFSS)
to examine the relationship between the density of park land and
physical activity and obesity rates for 67 metropolitan statistical
areas in the U.S. The findings in this case conformed to expectation;
that is, the study found a significant positive association between park
density and physical activity and a significant negative association
between park density and obesity. \citet{Larson2016} similarly used
self-reported scores on the Gallup-Healthways Wellbeing Index to
evaluate the relationship between physical health and park quantity,
quality, and accessibility in 44 U.S. cities. While the authors found
positive associations between wellbeing and park quality and
accessibility, these relationships were not statistically significant.
Conversely, \citet{Richardson2012} examined the relationship between
total urban greenspace and mortality rates for selected maladies; though
the authors did not find a statistically significant relationship
between the quantity of urban greenspace and mortality caused by lung
cancer, diabetes, heart disease or car accidents, they did find that
all-cause mortality was --- oddly --- \emph{higher} in greener cities.
Finally, in a meta-analysis of 20 peer reviewed journal articles
exploring the relationship between parks and objectively measured
physical activity, \citet{Bancroft2015} found that five studies reported
a significant positive association between the two, six studies produced
mixed results, and nine studies found no association at all.

Metropolitan-level analyses such as those described above do not capture
the within-city variation in accessibility to parks that exist in many
cities. Some cities with large amounts of total greenspace may
nevertheless unequally distribute this space throughout the city,
leading to areas with poor access. Conversely, a city with a smaller
overall proportion of greenspace may give all of its citizens better
access to sufficient greenspace to meet their needs or desires for
physical activity. Considering access at the neighborhood level within a
single city may also eliminate some regional or cultural fixed effects
affecting metropolitan-level analyses.

\hypertarget{park-accessibility-measures}{%
\subsection{Park Accessibility
Measures}\label{park-accessibility-measures}}

Loosely defined, the \emph{accessibility} of an arbitrary place
describes the ease with which people can accomplish activities there.
Accessibility is an abstract concept with tempting quantifiability
\citep{Handy1997}, perhaps explaining the proliferation of calculation
mechanisms. \citet{Dong2006} present a helpful mathematical heirarchy of
common mechanisms, which we follow here. Consider a person residing at
point \(i\) in a city with parks \(j \in 1 \ldots J\). An analyst might
consider point \(i\) as ``having access'' to a park if the distance to
the park \(d_{ij}\) is within an ``isochrone,'' or a geometric buffer
defined by some threshold \(D\),

@@

Variations of this isochrone-based framework are easily derived and
relatively common. ParkScore \citep{parkscore2019} is a sophisticated
application of this approach where \(D\) is a carefully calculated
10-minute network-bound walk, paired with a demographic analysis of
areas within and outside of this threshold. In other applications \(D\)
might be more variably defined, such as the presence or percent of green
space within a political or statistical boundary
\citep{Mitchell2008, Stark2014} or within a specific distance threshold
\citep{Kaczynski2014}. It is also possible to attenuate an isochrone
with various factors; for instance, \citet{Dias2019} considered road
safety perception as an extenuating factor in the effective distance to
a park.

In spite of the flexibility of adapted isochrone techniques, the
arbritrary definitions necessarily imposed by researchers in its
application can inhibit holistic analysis \citep{Logan2017}. A somewhat
more complete approach is the so-called ``gravity'' accessibility
statistic. In this case the accessibility of point \(i\) is the
denominator of the gravity formulation of a trip distribution model,
\begin{equation}\label{eq:gravity}
A_i = \sum_{j = 1}^J S_j f(d_{ij})
\end{equation} where \(S_j\) is the ``size'' of each park literally
(i.e.~acres) or abstractly (i.e. trip attractions) and \(f(d_{ij})\) is
a monotonically decreasing cost function, usually a negative
exponential. \citet{Dong2006} note that the isochrone framework is a
special case of the gravity model where \(f(d_{ij})\) is a binary
function and \(S_j = 1\). \citet{Giles-Corti2005} compare various
gravity-based accessibility scores with an isochrone specification and
show that the former are more predictive of walking behavior; that is,
the attributes of a park influence walking more than merely the distance
to it. \citet{Zhang2011} develop a population-weighted, gravity-based
accessibility to parks metric with a national scope, though they do not
examine its correlation with health outcomes.

A third accessibility mechanism is a utility-based specification, termed
as such for being derived from random utility choice theory. Consider
that an individual living at point \(i\) is choosing a park for a
recreation activity. If we apply the multinomial logit model
\citep{McFadden1974}, the expected consumer surplus enjoyed by this
individual can be shown to be \begin{equation}\label{eq:utility}
A_i = \ln\left({\sum_{j = 1}^J\exp(V_{ij})}\right) + C
\end{equation} where parks are differentiated from each other by their
relative measurable \emph{utilities}, \(V_{ij}\). \(C\) is an unknown
constant, but the difference in consumer surplus between two points
\(i\) and \(k\) can be quantitatively compared as \(A_i - A_k\)
\citep{Bruce1977}. In principle, \(V\) may include any measurable
attributes of either the choice maker or the park, and is typically
represented as a linear-in-parameters function of destination attributes
\(X_j\) and the travel cost \(d_{ij}\),
\begin{equation}\label{eq:utilityV}
V_{ij} = X_j\beta + \beta_d * d_{ij}
\end{equation} The coefficients \(\vect{\beta}\) are frequently
estimated from household surveys, though in the absence of a survey we
may assert reasonable values.

Note that the gravity formulation is itself a special case of the
utility-based specification where
\(\exp(X_j\beta + \beta_d d_{ij}) = X_j f(d_{ij})\) \citep{Daly1982}.
The distinction between gravity and utility-based specification is
meaningful, however. Primarily, it becomes possible to construct
accessibility statistics based on revealed behavioral preferences rather
than calibrated or asserted values \citep{Handy1997}.
\citet{Kaczynski2016} present what is effectively a utility-based
accessibility score they call ``ParkIndex'', with the explicit
motivation of developing a uniformly applicable park accessibility
statistic, though they limit this statistic as a way to add
heterogeneity within an isochrone analysis.

@@

To this dataset of parks and regularized green spaces, we add a polygons
layer of cemeteries in New York that are open to the public. Cemeteries
are important green spaces that can be used for many types of physical
activity. These operations leave us with 1,277 distinct green spaces;
for simplicity we will refer to both parks and cemeteries as ``parks''
going forward. For each park, we calculate the size of the park in
acres. Finally, using a Twitter application programming interface (API)
implemented with the Python package Tweepy, we also collected and stored
tweets containing geotags of precise geographic coordinates located
within the boundaries of the parks; from this information, we were able
to segregate tweets that were generated within a park in September 2014.
This geolocated twitter activity could provide an additional point of
information on the degree to which individual parks are actually used
\citep{Wang2016}.

@@

\caption{\label{tab:tractsdata}Descriptive Statistics of Tract and Park Variables}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{l>{\raggedright\arraybackslash}p{2in}rlrl}
\toprule
  & Description & Minimum & Median (IQR) & Maximum & Source\\
\midrule
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Tract Variables, N = 2102}}\\
\hspace{1em}Obesity & Share of adults over 18 who are obese & 10.20 & \makecell{24.90\\ (19.83, 30.80)} & 45.40 & CDC 500 Cities\\
\hspace{1em}Physical Activity & Share of adults over 18 who engage in some leisure-time physical activity & 45.90 & \makecell{72.10\\ (66.60, 76.77)} & 90.70 & CDC 500 Cities\\
\hspace{1em}Income & Median tract income & 9053.00 & \makecell{59,592.50\\ (41,928.25, 79,092.75)} & 250001.00 & ACS\\
\hspace{1em}Density & Households per square kilometer & 9.20 & \makecell{5,848.46\\ (3,173.62, 9,674.36)} & 43621.52 & ACS\\
\hspace{1em}Fulltime & Share of adults over 18 with full-time work & 8.80 & \makecell{49.48\\ (44.45, 55.29)} & 100.00 & ACS\\
\hspace{1em}College & Share of adults over 24 with a college degree & 0.61 & \makecell{16.30\\ (12.37, 20.11)} & 44.94 & ACS\\
\hspace{1em}Single & Share of adults over 18 living alone or in a non-partnership household & 16.38 & \makecell{59.39\\ (50.30, 68.65)} & 100.00 & ACS\\
\hspace{1em}Youth & Share of population under 18 & 0.00 & \makecell{20.54\\ (16.79, 24.92)} & 64.07 & ACS\\
\hspace{1em}Young adults & Share of population between 18 and 34 & 0.00 & \makecell{25.73\\ (21.66, 29.98)} & 86.75 & ACS\\
\hspace{1em}Seniors & Share of population who are 65 and over & 0.00 & \makecell{12.83\\ (9.47, 16.88)} & 89.88 & ACS\\
\hspace{1em}Black & Share of population who is black & 0.00 & \makecell{10.03\\ (2.13, 44.62)} & 220.65 & ACS\\
\hspace{1em}Asian & Share of population who is Asian & 0.00 & \makecell{7.66\\ (2.40, 20.78)} & 88.07 & ACS\\
\hspace{1em}Hispanic & Share of population who is Hispanic & 0.00 & \makecell{19.07\\ (9.39, 41.07)} & 96.27 & ACS\\
\hspace{1em}Other & Share of population who belong to other minority groups & 0.00 & \makecell{0.00\\ (0.00, 0.53)} & 19.47 & ACS\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Park Variables, N = 1277}}\\
\hspace{1em}Size & Park size in acres & 0.50 & \makecell{1.66\\ (0.95, 5.62)} & 1961.00 & NYC\\
\hspace{1em}Courts & Presence of sport courts / ball fields & 0.00 & \makecell{0.00\\ (0.00, 0.00)} & 1.00 & NYC\\
\hspace{1em}Playgrounds & Presence of playgrounds & 0.00 & \makecell{0.00\\ (0.00, 1.00)} & 1.00 & NYC\\
\hspace{1em}Trails & Presence of trails & 0.00 & \makecell{0.00\\ (0.00, 0.00)} & 1.00 & NYC\\
\hspace{1em}Tweets & Tweets emanating from park in September 2014 & 0.00 & \makecell{0.00\\ (0.00, 2.00)} & 1593.00 & Twitter API\\
\bottomrule
\end{tabular}}
\end{table*}

@@

We calculate a set of utility-based accessibility statistics for each
tract in New York City. The most basic utility specification includes
only the park size in acres and the distance from the
population-weighted tract centroid to the boundary of the park in miles,
\begin{equation}\label{eq:u_access}
V_{ij} = \lambda_s * \log(size_j) + \lambda_d * \log(distance_{ij})
\end{equation} The logarithmic transform allows for diminishing marginal
utility of distance and park size: A 1-mile increase to a trip matters
more for a 1-mile trip than a 10-mile trip. \citet{Macfarlane2019}
estimated destination choice parameters for park trips in Alameda County
(Oakland), California by applying this utility specification to
passively collected mobile device data, obtaining values of
\(\lambda_s = 0.373\) and \(\lambda_d = -1.76\). The ratio of these
estimates implies people are willing to travel roughly six times further
to reach a park twice as large. We do not have access to the necessary
data to repeat the \citet{Macfarlane2019} methodology in New York, and
the public parks datasets in Alameda County are not sufficiently
detailed to perform the present accessibility analysis there. In the
absence of park trip destination choice coefficients in New York City,
we adopt these previous estimates.

@@

The third utility specification includes the number of geolocated tweets
emanating from within the park in September of 2014 in addition to the
size and distance terms in Equation \ref{eq:u_access},
\begin{equation}\label{eq:u_tweets}
\begin{split}
V_{ij} &= \lambda_s * \log(size_j) + \lambda_d * \log(distance_{ij}) + \\ 
  &\quad \lambda_{tw} * YJ(tweets)
\end{split}
\end{equation} with \(YJ(x)\) being a Yeo-Johnson transformation that
implements a diminishing marginal return similar to \(log(x)\), but
where \(YJ(0) = 0\) \citep{Yeo2000}. In this case we have no external
information describing how destination choice to parks may be affected
by the twitter activity in the park, so we assert a value of
\(\lambda_{tw} = 0.1\)

We then calculate the utility-based accessibility of each tract \(A_i\)
as defined in Equation \ref{eq:utility} under each utility
specification. Recall that the total value of the accessibility is
relative to an unknown scalar \(C\); for this reason we standardize the
accessibility values for all tracts within each utility specification,
\begin{equation}\label{eq:v_logged}
A_i' = \frac{A_i - \bar{A}}{sd(A)}
\end{equation}

@@

We predict either the physical activity or the obesity rate (\(y\)) in a
census tract as a linear function of the tract's sociodemographic
characteristics \(X\) and accessibility to parks \(A\),
\begin{equation}\label{eq:themodel}
 y = X\beta + \beta_{a}A + \varepsilon
\end{equation}

As the observations are related to each other spatially, it is likely
that this process involves spatial spillovers of one kind or another. A
complete treatment of spatial econometrics is not warranted here; the
reader is directed to \citet{LeSage2009} as well as \citet{LeSage2014}.
Suffice it to say that the presence of spatial data generating processes
can negatively affect econometric interpretation in a variety of ways.
Each of these processes rely on spatial autoregression, where elements
of some random variable \(x\) are spatially dependent on other elements,
\begin{equation}\label{eq:autoregression}
 x = \rho W x + \varepsilon
\end{equation} with the strength of dependence an estimated parameter
\(\rho\), the structure of spatial dependence governed by the asserted
matrix \(W\) (details of \(W\) follow below), and some random
uncorrelated residual \(\varepsilon\). A collection of models is
available to represent spatial autodependence in the model residuals,
the independent variables, and the dependent variable.

The specific spatial model applied is both an econometric choice --- as
improperly specified spatial models can lead to inconsistent estimates
of model parameters and standard errors --- as well as a philosophical
choice about the likely structure of spillovers in the problem at hand.
In case of physical activity and obesity rates, we believe that the
outcomes are locally dependent: That is, a particular individual decides
whether or not to participate in physical activity independently of
whether his or her neighbors participate in physical activity. This does
not mean that we think it impossible that the socioeconomic variables of
neighboring tracts do not play a role. Mathematically, we assert that
the autodependence relationship on the dependent variable
\(y = \rho W y + \ldots\) has \(\rho = 0\) in the current case.

>>>>

\documentclass[shortAfour,sageh.bst]{sagej}

@@

\usepackage{bm}
     \def\Var{{\rm Var}\,}
     \def\Cov{{\rm Cov}\,}
     \def\mean{{\rm mean}\,}
     \def\E{{\rm E}\,}
  	 \def\Corr{{\rm Corr}\,}

\usepackage{xcolor}

\usepackage{hyperref}
\usepackage{rotating}
\usepackage{booktabs}
\usepackage{makecell}
\usepackage{dcolumn}
\newcolumntype{d}[1]{D{.}{.}{#1}}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

\runninghead{Macfarlane \emph{et al}.}

\author{Gregory Macfarlane\affilnum{1}, Nico Boyd\affilnum{2}, John E. Taylor\affilnum{3}, Kari Watkins\affilnum{3}}

\affiliation{\affilnum{1}{Department of Civil \& Environmental Engineering, Brigham Young University}\\\affilnum{2}{Fehr \& Peers, Los Angeles California}\\\affilnum{3}{School of Civil \& Environmental Engineering, Georgia Institute of Technology}}

\corrauth{Gregory Macfarlane, 430 Engineering Building, Provo UT 84663}

\email{\href{mailto:gregmacfarlane@byu.edu}{\nolinkurl{gregmacfarlane@byu.edu}}}

@@

\keywords{Accessibility; Green Space}

\maketitle
\clearpage
\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

@@

Historically, most large-scale health promotion efforts focused on
individual-level interventions intended to educate people about healthy
lifestyles and behaviors, touching on topics including diet and
exercise. Over the last several years a new \emph{social} model of
health has evolved, described by \citet{Duhl1999} ``as an outcome of the
effects of socioeconomic status, culture, environmental conditions,
housing, employment and community
influences''\citeyearpar[p.7]{Duhl1999}. In this new paradigm, resources
provided via civil infrastructure --- in particular parks and other
public green spaces --- play a crucial role in promoting and sustaining
public health \citep{Bedimo-Rung2005, Wells2007, Coutts2008}. After all,
it does little to educate individuals on the importance of exercise if
their community is designed and built in such a way that such exercise
is expensive, unenjoyable, or unsafe. It is similarly necessary for
urban planners to identify which neighborhoods have access to quality
green space, and which neighborhoods need intervention in the form of
new or rehabilitated green space, or improved transportation access to
existing quality green space.

But in spite of the potential for green spaces to influence public
health outcomes and numerous studies attempting to empirically estimate
the strength of these outcomes, evidence to date has been mixed. A major
challenge researchers have faced is a proliferation of techniques for
measuring access to green spaces, of which many are plausible but lack
robust theoretical bases.

In this paper, we apply a holistic and flexible measurement for park
accessibility based in activity location choice theory. This
utility-based accessibility measure compares the continuous distance to
all parks in the region, weighted against the sizes of each park and its
other assorted amenities. We apply this measure to study the link
between park accessibility and attractiveness and census tract-level
aggregate physical activity participation and obesity rates in New York
City, controlling for spatially correlated unobserved effects, spatial
spillovers of covariates, and tract-level socioeconomic characteristics.
We also demonstrate how the utility-based accessibility metric can
potentially be expanded to account for other measures of a park's
attractiveness in the form of various park amenities. The paper
concludes with a discussion of opportunities for future research.

@@

\label{sec:litreview} The question of how green space in the urban
environment affects the physical, emotional, and mental well-being of
urban residents is not new to the academic literature. Comprehensive
reviews and analysis of this literature up to 2007 can be had from
\citet{tzoulas2007promoting}, and for the subsequent decade from
\citet{kabisch2015human}. In spite of the depth of this research arena
--- or perhaps because of it --- quantitative approaches for assessing
both health outcomes and access to green space vary widely. It is
therefore not surprising that in a meta-analysis of 20 recent peer
reviewed journal articles exploring the relationship between access to
parks and objectively measured physical activity, \citet{Bancroft2015}
found that five studies reported a significant positive association, six
studies produced mixed results, and nine studies found no association at
all.

Loosely defined, the \emph{accessibility} of an arbitrary place
describes the ease with which people can accomplish activities there.
Accessibility is an abstract concept with tempting quantifiability
\citep{Handy1997}, but objective quantification may not be strictly
necessary. For example, \citet{takano2002urban} showed improved survival
rates among senior citizens in Japan who self-reported having access to
qualitatively good walking spaces. Though encouraging, this result could
simply reveal self-selection: Those seniors who felt the walking space
was good were those who chose to enjoy it. Urban planners seeking to
evaluate which residents have good access to quality green space may
desire objective and therefore comparable measures.

\citet{Dong2006} present a helpful mathematical hierarchy of common
quantitative mechanisms for calculating access, which we follow here.
Consider a person residing at point \(i\) in a city with parks
\(j \in 1 \ldots J\). An analyst might consider point \(i\) as ``having
access'' to a park if the distance to the park \(d_{ij}\) is within an
``isochrone,'' or a geometric buffer defined by some threshold \(D\),

@@

Applications of this isochrone-based framework are easily derived and
relatively common. ParkScore \citep{parkscore2019} is a sophisticated
application of this approach where \(D\) is a carefully calculated
10-minute network-bound walk, paired with a demographic analysis of
areas within and outside of this threshold. \citet{devries2003natural}
showed the amount of greenspace within a 3-kilometer radius was
positively associated with a suite of self-reported health outcomes.
Similarly, \citet{devries2013streetscape} showed that streetscape
greenery within 500 meters of a residence was associated with perceived
general health. Conversely, \citet{carlson2012interactions} found no
relationship between the health of senior citizens and the number of
parks within 500 meters.

Another common access measure fits in this category, where
distance-based isochrones are replaced with political, geographical, or
statistical boundaries. For example, \citet{Mitchell2008} find that
inequalities in health outcomes attributable to income are mitigated in
statistical areas with a greater percentage of green space. These are
especially common in comparative macroscopic studies, such as the
positive relationships found by \citet{West2012} between density of park
land in a metropolitan area and self-reported population health
indicators derived from the Behavioral Risk Factor Surveillance System
(BRFSS). On the other hand, \citet{Larson2016} found no relationship
between physical health and park quantity, quality, and accessibility in
44 U.S. cities and self-reported scores on the Gallup-Healthways
Wellbeing Index. At the sub-metropolitan level, \citet{Stark2014} found
a significant relationship in New York City between the percent of a
residential ZIP code that is park space and objectively measured body
mass index (BMI). It is also possible to attenuate an isochrone with
various factors; for instance, \citet{Dias2019} considered road safety
perception as an extenuating factor in the effective distance to a park.

In spite of the flexibility of adapted isochrone techniques, the
arbitrary definitions necessarily imposed by researchers in its
application can inhibit holistic analysis \citep{Logan2017}. A somewhat
more complete approach is the so-called ``gravity'' accessibility
statistic. In this case the accessibility of point \(i\) is the
denominator of the gravity formulation of a trip distribution model,
\begin{equation}\label{eq:gravity}
A_i = \sum_{j = 1}^J S_j f(d_{ij})
\end{equation} where \(S_j\) is the ``size'' of each park literally
(i.e.~acres) or abstractly (i.e. trip attractions) and \(f(d_{ij})\) is
a monotonically decreasing cost function, usually a negative
exponential. \citet{Dong2006} note that the isochrone framework is a
special case of the gravity model where \(f(d_{ij})\) is a binary
function and \(S_j = 1\). \citet{Giles-Corti2005} compare various
gravity-based accessibility scores with an isochrone specification and
show that the former are more predictive of walking behavior; that is,
the attributes of a park influence walking more than merely the distance
to it. \citet{hillsdon2006relationship} found no relationship between a
gravity-based accessibility term and self-reported physical activity
among middle-aged individuals in Norwich, England. \citet{Zhang2011}
developed a population-weighted, gravity-based accessibility to parks
metric with a national scope, though they do not examine its correlation
with health outcomes.

A third accessibility mechanism is a utility-based specification, termed
as such for being derived from random utility choice theory. Consider
that an individual living at point \(i\) is choosing a park for a
recreation activity. If we apply the multinomial logit model
\citep{McFadden1974}, the expected consumer surplus enjoyed by this
individual can be shown to be \begin{equation}\label{eq:utility}
A_i = \ln\left({\sum_{j = 1}^J\exp(V_{ij})}\right) + C
\end{equation} where parks are differentiated from each other by their
relative measurable \emph{utilities}, \(V_{ij}\). \(C\) is an unknown
constant, but the difference in consumer surplus between two points
\(i\) and \(k\) can be quantitatively compared as \(A_i - A_k\)
\citep{Bruce1977}. In principle, \(V\) may include any measurable
attributes of either the choice maker or the park, and is typically
represented as a linear-in-parameters function of destination attributes
\(X_j\) and the travel cost \(d_{ij}\),
\begin{equation}\label{eq:utilityV}
V_{ij} = X_j\boldsymbol{\beta} + \beta_d * d_{ij}
\end{equation} The coefficients \(\boldsymbol{\beta}\) are frequently
estimated from household surveys, though in the absence of a survey we
may assert reasonable values.

Note that the gravity formulation is itself a special case of the
utility-based specification where
\(\exp(X_j\boldsymbol{\beta} + \beta_d d_{ij}) = X_j f(d_{ij})\)
\citep{Daly1982}. The distinction between gravity and utility-based
specification is meaningful, however. Primarily, it becomes possible to
construct accessibility statistics based on revealed behavioral
preferences rather than calibrated or asserted values \citep{Handy1997}.
\citet{Kinnell2006} collected a choice-based survey sufficient to
construct a utility-based measure, though they do not do so. Similarly,
\citet{Kaczynski2016} present what is effectively a utility-based score
called ``ParkIndex'', with the explicit motivation of developing a
uniformly applicable park accessibility statistic, though they limit
this statistic as a way to add heterogeneity within an isochrone
analysis.

@@

To this dataset of parks and regularized green spaces, we add a polygons
layer of cemeteries in New York that are open to the public. Cemeteries
are important green spaces that can be used for many types of physical
activity. These operations leave us with 1,277 distinct green spaces;
for simplicity we will refer to both parks and cemeteries as ``parks''
going forward. For each park, we calculate the size of the park in
acres.

@@

\caption{\label{tab:tractsdata}Descriptive Statistics of Tract and Park Variables}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{l>{\raggedright\arraybackslash}p{2in}rlrl}
\toprule
  & Description & Minimum & Median (IQR) & Maximum & Source\\
\midrule
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Tract Variables, N = 2102}}\\
\hspace{1em}Obesity & Share of adults over 18 who are obese & 9.60 & \makecell{24.50\\ (19.60, 30.50)} & 44.20 & CDC 500 Cities\\
\hspace{1em}Physical Activity & Share of adults over 18 who engage in some leisure-time physical activity & 48.10 & \makecell{69.50\\ (64.70, 74.60)} & 89.10 & CDC 500 Cities\\
\hspace{1em}Income & Median tract income & 9053.00 & \makecell{59,592.50\\ (41,928.25, 79,092.75)} & 250001.00 & ACS\\
\hspace{1em}Density & Households per square kilometer & 9.20 & \makecell{5,848.46\\ (3,173.62, 9,674.36)} & 43621.52 & ACS\\
\hspace{1em}Fulltime & Share of adults over 18 with full-time work & 8.80 & \makecell{49.48\\ (44.45, 55.29)} & 100.00 & ACS\\
\hspace{1em}College & Share of adults over 24 with a college degree & 0.61 & \makecell{16.30\\ (12.37, 20.11)} & 44.94 & ACS\\
\hspace{1em}Single & Share of adults over 18 living alone or in a non-partnership household & 16.38 & \makecell{59.39\\ (50.30, 68.65)} & 100.00 & ACS\\
\hspace{1em}Youth & Share of population under 18 & 0.00 & \makecell{20.54\\ (16.79, 24.92)} & 64.07 & ACS\\
\hspace{1em}Young adults & Share of population between 18 and 34 & 0.00 & \makecell{25.73\\ (21.66, 29.98)} & 86.75 & ACS\\
\hspace{1em}Seniors & Share of population who are 65 and over & 0.00 & \makecell{12.83\\ (9.47, 16.88)} & 89.88 & ACS\\
\hspace{1em}Black & Share of population who is black & 0.00 & \makecell{10.03\\ (2.13, 44.62)} & 220.65 & ACS\\
\hspace{1em}Asian & Share of population who is Asian & 0.00 & \makecell{7.66\\ (2.40, 20.78)} & 88.07 & ACS\\
\hspace{1em}Hispanic & Share of population who is Hispanic & 0.00 & \makecell{19.07\\ (9.39, 41.07)} & 96.27 & ACS\\
\hspace{1em}Other & Share of population who belong to other minority groups & 0.00 & \makecell{0.00\\ (0.00, 0.53)} & 19.47 & ACS\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Park Variables, N = 1277}}\\
\hspace{1em}Size & Park size in acres & 0.50 & \makecell{1.66\\ (0.95, 5.62)} & 1961.00 & NYC\\
\hspace{1em}Courts & Presence of sport courts / ball fields & 0.00 & \makecell{0.00\\ (0.00, 0.00)} & 1.00 & NYC\\
\hspace{1em}Playgrounds & Presence of playgrounds & 0.00 & \makecell{0.00\\ (0.00, 1.00)} & 1.00 & NYC\\
\hspace{1em}Trails & Presence of trails & 0.00 & \makecell{0.00\\ (0.00, 0.00)} & 1.00 & NYC\\
\bottomrule
\end{tabular}}
\end{table*}

@@

We calculate a pair of utility-based accessibility statistics for each
tract in New York City. The most basic utility specification includes
only the park size in acres and the Euclidean distance from the
population-weighted tract centroid to the nearest boundary of the park
in miles, \begin{equation}\label{eq:u_access}
V_{ij} = \lambda_s * \log(size_j) + \lambda_d * \log(distance_{ij})
\end{equation} There are a few cases where the population-weighted tract
centroid lies on or within the boundary of a park; for these cases we
assert that the minimum possible distance is one tenth of a mile. The
logarithmic transform allows for diminishing marginal utility of
distance and park size: A 1-mile increase to a trip matters more for a
1-mile trip than a 10-mile trip. \citet{Macfarlane2019} estimated
destination choice parameters for park trips in Alameda County
(Oakland), California by applying this utility specification to
passively collected mobile device data, obtaining values of
\(\lambda_s = 0.373\) and \(\lambda_d = -1.76\). The ratio of these
estimates implies people are willing to travel roughly six times further
to reach a park twice as large. We do not have access to the necessary
data to repeat the \citet{Macfarlane2019} methodology in New York, and
the public parks datasets in Alameda County are not sufficiently
detailed to perform the present accessibility analysis there. In the
absence of park trip destination choice coefficients in New York City,
we adopt these previous estimates.

@@

We then calculate the utility-based accessibility of each tract \(A_i\)
as defined in Equation \ref{eq:utility} under each utility
specification. Recall that the total value of the accessibility is
relative to an unknown scalar \(C\); for this reason we standardize the
accessibility values for all tracts within each utility specification,
\begin{equation}\label{eq:v_logged}
A_i' = \frac{A_i - \bar{\boldsymbol{A}}}
            {\sigma_{\boldsymbol{A}}}
\end{equation}

@@

We predict either the physical activity or the obesity rate (\(y\)) in a
census tract as a linear function of the tract's sociodemographic
characteristics \(X\) and accessibility to parks \(\boldsymbol{A}\),
\begin{equation}\label{eq:themodel}
 \boldsymbol{y} = X\boldsymbol{\beta} + \beta_{a}\boldsymbol{A} + \boldsymbol{\varepsilon}
\end{equation} where \(X\) is a matrix of dimension \(n\times p\) where
\(n\) is the number of observations and \(p\) is the number of predictor
variables, and \(\boldsymbol{\beta}\) is a \(p\times 1\) vector of
estimated parameters. The accessibility vector \(\boldsymbol{A}\)
influences physical activity or obesity rates by the single estimated
parameter \(\beta_a\), and a vector \(\boldsymbol{\varepsilon}\)
represents the model residual as a random variable.

As the observations are related to each other spatially, it is likely
that this process involves spatial spillovers of one kind or another. A
complete treatment of spatial econometrics is not warranted here; the
reader is directed to \citet{LeSage2009} as well as \citet{LeSage2014}.
Suffice it to say that the presence of spatial data generating processes
can negatively affect econometric interpretation in a variety of ways.
Each of these processes rely on spatial autoregression, where elements
of some random variable \(x\) are spatially dependent on other elements,
\begin{equation}\label{eq:autoregression}
 \boldsymbol{x} = \rho W \boldsymbol{x} + \boldsymbol{\varepsilon}
\end{equation} with the strength of dependence an estimated parameter
\(\rho\), the structure of spatial dependence governed by the asserted
matrix \(W\) (details of \(W\) follow below), and some random
uncorrelated residual \(\varepsilon\). A collection of models is
available to represent spatial autodependence in the model residuals,
the independent variables, and the dependent variable.

The specific spatial model applied is both an econometric choice --- as
improperly specified spatial models can lead to inconsistent estimates
of model parameters and standard errors --- as well as a philosophical
choice about the likely structure of spillovers in the problem at hand.
In case of physical activity and obesity rates, we believe that the
outcomes are locally dependent: That is, a particular individual decides
whether or not to participate in physical activity independently of
whether his or her neighbors participate in physical activity. This does
not mean that we think it impossible that the socioeconomic variables of
neighboring tracts do not play a role. Mathematically, we assert that
the autodependence relationship on the dependent variable
\(\boldsymbol{y} = \rho W \boldsymbol{y} + \ldots\) has \(\rho = 0\) in
the current case.
